---
title: "re_stepwise"
author: "Fabio Jakob 20-107-876"
date: "2023-04-17"
output: html_document
---
read the data with readr
```{r}
library(readr)
df_for_stepwise_regression <- read_csv('C:/Users/fabio.LAPTOP-FR4VQFHF/Documents/agds_report_fabiojakob/Data/df_for_stepwise_regression.csv')
View(df_for_stepwise_regression)
```
1. Evaluation of all bivariate models

implementing steps 1-3 of the algorithm.

Define the response variable and predictor variables
```{r}
response <- "GPP_NT_VUT_REF"
predictors <- c("siteid", "TIMESTAMP", "TA_F", "SW_IN_F", "LW_IN_F", "VPD_F", "PA_F", "P_F", "WS_F", "TA_F_MDS", "SW_IN_F_MDS", "LW_IN_F_MDS", "VPD_F_MDS", "CO2_F_MDS", "PPFD_IN", "USTAR")
```

Initialize an empty vector to store the selected predictors
```{r}
selected <- c()
```


Set the significance level for adding predictors to the model
```{r}
alpha <- 0.05
```

Fit a linear regression model with each predictor added one at a time
```{r}
p <- rep(NA, length(predictors))
  
for (i in 1:length(predictors)) {
  formula <- paste(response, paste(c(selected, predictors[i]), collapse = "+"), sep = "~")
  model <- lm(formula, df_for_stepwise_regression)
  p[i] <- summary(model)$coef[2, 4]
}
```

Find the predictor with the lowest p-value
```{r}

min_p <- min(p)
min_index <- which(p == min_p)
best_predictor <- predictors[min_index]
```


To visualize the results we can create a bar plot who shows the p-value of every model
```{r}

results <- data.frame(predictors, p)


library(ggplot2)
ggplot(results, aes(x = predictors, y = p)) + 
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "", y = "P-Value") +
  ggtitle("P-Values of Predictors") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(trans = "log10")
```
Because the some p-values are very small i tried to fit results in a logarithmic scale but some values are still not visible because they are too small. We can still see which variables have a high p-value and therefore are not suitable for a final subset of predictors.

2. manual implementation of stepwise forward regression:

We use the same code as above and put it into a loop to add the suitable predictors one after another
```{r}
while (length(predictors) > 0) {
  # Initialize a vector to store the p-values for each predictor
  p <- rep(NA, length(predictors))
  
  # Fit a linear regression model with each predictor added one at a time
  for (i in 1:length(predictors)) {
    formula <- paste(response, paste(c(selected, predictors[i]), collapse = "+"), sep = "~")
    model <- lm(formula, df_for_stepwise_regression)
    p[i] <- summary(model)$coef[2, 4]
  }
  
  # Find the predictor with the lowest p-value
  min_p <- min(p)
  min_index <- which(p == min_p)
  best_predictor <- predictors[min_index]
  
  # Add the best predictor to the selected list
  selected <- c(selected, best_predictor)
  
  # Remove the best predictor from the list of predictors
  predictors <- predictors[-min_index]
  
  # Check if the best predictor is significant enough to keep in the model
  if (min_p < alpha) {
    cat(paste("Added", best_predictor, "to the model.\n"))
  } else {
    cat(paste("Stopped the algorithm. No more predictors are significant.\n"))
    break
  }
}
```

Fit a final linear regression model with the selected predictors
```{r}
formula <- paste(response, paste(selected, collapse = "+"), sep = "~")
model <- lm(formula, df_for_stepwise_regression)
```

Print the final model summary
```{r}
summary(model)
```

If we look at the R-squared we see that the value is 0.5884. This is not a very bad value but there are
sure better options for a better model. But it is questionable if a higher R-squared is possible with the 
methods that we've learned so far and the fact that the data set is not the easiest doesn't help.
To visualize ho well the model fits the data we can predict the values of the response variable with the model using the predict function and put this values in a new column of the data set. Then we can create a scatter plot to see how the predicted and the actual values correlate.
```{r}
df_for_stepwise_regression$y_pred <- predict(model, newdata = df_for_stepwise_regression)

library(ggplot2)
ggplot(df_for_stepwise_regression, aes(x = GPP_NT_VUT_REF, y = y_pred)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Actual Values", y = "Predicted Values", title = "Stepwise Forward Regression Results")
```
We see that the correlation is not that bad but also not perfect and the values are at least ïn the same magnitude.
















